Automated iterative forecasting for the Portal Project

### 2018-02-02 manuscript/portal_iterative_forecasting_ms.Rmd
Revise changes to modeling and forecasting section
This section had been overly softened in ways that make things less clear.

Iterative near-term forecasting involves regularly refitting a variety of different models (Figure 1). Additionally, any new models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach. We use CI to refit the models and make new forecasts each time the modeling code changes and when new data becomes available (Figure 2). We use a plugin infrastructure to allow new models to be easily added to the system. Our approach is to treat each model as interchangable black boxes - we feed them all the same inputs and we expect them to generate the same structure for model outputs (Figure 3). Each iteration of the forecasting cycle, all models are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts. This structure allows weighted ensemble models to be produced based on how well individual models fit the training data. Our hope is that by requiring all models to take the same input (though models may differ in how they use that input) and provide the same outputs, we've gained the capacity to add and compare very different types of models that we may not currently be able to foresee and design around. Our infrastrcture should allow easy addition of more complicated models and more sophisticated approaches to creating ensembles even though to date our models and ensemble-creating approaches are extremely basic.

@@

We designed our forecast output format to to store relatively generic forecasts information to allow flexibility in what the models predict (i.e. species abundances, biomass). Each forecast output file contains the date being forecast, the collection date of the data used for fitting the models, the date the forecast was made, the state variable being forecast (e.g., rodent biomass, the abundance of a species), and the forecast value and associated uncertainty of that forecast (Figure 3). Our goal is to store a variety of different forecasts in a common format with the hope of this being a useful starting point for developing a standard for storing ecological forecasts more generally.

Our forecasts are currently evaluated using root mean square error (RMSE) to evaluate point forecasts and coverage to evaluate uncertainty. We plan to add additional metrics in the future. In addition to evaluating the actual forecasts, we also use hindcasting (forecasting on already collected data) to gain additional insight into the methods that work best for forecasting this system. For example, a model is fit using rodent observations up to June 2005, then used to make a forecast 12 months out to May 2006. The observations of that 12 month period can immediately be used to evaluate the model. Hindcasting can be conducted using data that was already been collected, thus allowing model comparison of large numbers of hindcasts and giving insight into which models make the best forecasts without needing to wait for new data to be collected. It can also be used to quickly evaluate new models instead of waiting for an adequate amount of data to accumulate.

>>>>

Iterative near-term forecasting involves regularly refitting a variety of different models (Figure 1). Ideally, new models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach. We use CI to refit the models and make new forecasts each time the modeling code changes and when new data becomes available (Figure 2). We use a plugin infrastructure to allow new models to be easily added to the system. This approach treats each model as an interchangable black boxes - all models have access to the same input data and generate the same structure for model outputs (Figure 3). Each iteration of the forecasting cycle, all models that are present are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts. A weighted ensemble model is then added with weights based on how well individual models fit the training data. This plugin infrastructure makes it easy to add and compare very different types of models, from the basic time-series approaches currently implemented to complex state-space and machine learning models. As long as a model script can load the provided data and produce the appropriate output it will be run and its results incorporated into the rest of the forecasting system.

@@

In addition to flexibility in model internals, we also wanted to support flexibility in what the models predict. Allowing models to make forecasts for system properties ranging from individual species' population abundances to total community biomass facilitates exploration of differences in forecastability across different aspects of ecological systems. We designed a forecast output format to support this. Each forecast output file contains the date being forecast, the collection date of the data used for fitting the models, the date the forecast was made, the state variable being forecast (e.g., rodent biomass, the abundance of a species), and the forecast value and associated uncertainty of that forecast (Figure 3). This allows us to store a variety of different forecasts in a common format and may serve as a useful starting point for developing a standard for storing ecological forecasts more generally.

Forecasts are currently evaluated using root mean square error (RMSE) to evaluate point forecasts and coverage to evaluate uncertainty. We plan to add additional metrics in the future. In addition to evaluating the actual forecasts, we also use hindcasting (forecasting on already collected data) to gain additional insight into the methods that work best for forecasting this system. For example, a model is fit using rodent observations up to June 2005, then used to make a forecast 12 months out to May 2006. The observations of that 12 month period can immediately be used to evaluate the model. Hindcasting can be conducted using data that was already been collected, thus allowing model comparison of large numbers of hindcasts and giving insight into which models make the best forecasts without needing to wait for new data to be collected. It can also be used to quickly evaluate new models instead of waiting for an adequate amount of data to accumulate.
